### 2015-08-15 text/ms.Rmd
Edits to ms

<!-- One solution is to... TODO KL: As part of setting the stage, right off the bat I'd delimit the set of conditions that make ensemble models an appropriate and useful choice, e.g.: When there are several models (how many to make it useful?) and there is a way to validate them, one solution is..."-->

@@

Here, we estimate population status and trajectory of exploited fish populations using ensembles and superensembles (collectively 'ensemble methods') of these four data-limited models. We apply a variety of approaches to both simulated and real-world fish stocks and compare their predictive performance. *TODO Add a few sentences on the rationale for how the approaches were identified and the testing was designed.*

<!--We find that ensembles, and in particular superensembles, generally improved predictive performance of status and trajectory of fish populations over any single model---especially when considered across multiple dimensions of predictive ability. Furthermore, the output from superensembles can aid mechanistic understanding of how individual models of status and trajectory perform.-->

@@

<!--
We tested the ability of ensemble models to improve estimates of population
status and trajectory when applied to both a large simulated dataset and a
global database of assessed stock status. Here we describe the datasets,
individual models of population status, and ensemble methods to combine those
estimates. We then describe how we evaluated the ability of the various models
to estimate population status.
-->

<!--
## Datasets
-->

We first developed and tested ensemble methods on a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulated data included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of biomass depletion at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamic settings: (1) a constant harvest rate, (2) a harvest rate coupled with biomass to mimic an open-access single-species fishery, (3) a scenario where harvest rate increases continuously, and (4) a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error. 
<!--TODO: make last sentence clearer
4. First paragraph in Methods describing simulations line 113 in pdf - you may want to describe the recrutiment variability a bit more. I know its not an important detail, but because its in the first paragraph of methods, it might turn people off reading further. Most people wouldn't understand what 'multiplcative' (including me) is or may not know what 'process error/noise' refers to. 
5. First paragraph in Methods last sentence - needed a bit of rewording because as was, i got quite confused. hopefully this new sentence makes sense and doesnt change the meaning. 
TODO: Jamie: add table instead / in addition to this paragraph?
-->

@rosenberg2014 ran ten iterations for each combination of factors adding stochastic draws of recruitment and catch-recording variability each time to generate a total of 5760 stocks. The simulation models were built in the package `FLR` [@kell2007] for the statistical software \textsf{R} [@r2015]. The code to generate the simulations is available at <https://github.com/flr/StockSims> (TODO: not there).

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database [@ricard2012] --- a compilation of stock-assessment output from hundreds of exploited marine populations around the world. Our analysis of the stock-assessment database was based on version XX downloaded on XX (TODO). After removing stocks for which the models described below did not converge (TODO: add number), this database included `r ram_stocks_n` stocks.
<!-- Dan: Do we address the implications of only running our method on stocks that converged later on? i.e. is there something about those omitted stocks that would make them less suited for this type of analysis?-->

@@

\noindent 
where $L$ is a generalized loss function, $f$ is some generalized regression function, $\hat{b}_\mathrm{CMSY}$ is the mean or slope prediction from the CMSY model etc.\ (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the spectral density of the scaled catch series at frequencies of 0.20 and 0.05, $b$ is the mean or slope value the model is fitted to, and $\theta$ is the set of parameters for the regression function. For all superensemble models of mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero --- we fit the models in log space and exponentiated the predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit superensemble models on the natural untransformed scale.
<!--add theta-hat description, b hat, make sure these are right, maybe simplify, describe arg min-->

>>>>

@@

Here, we estimate population status and trajectory of exploited fish populations using ensembles and superensembles (collectively 'ensemble methods') of these four data-limited models. We apply four ensemble-method approaches of increasing complexity to both simulated and real-world fish stocks and compare their predictive performance against each other and against individual model productions.

@@

<!--
We tested the ability of ensemble methods to improve estimates of population status and trajectory when applied to both a large simulated dataset and a global database of assessed stock status. Here we describe the datasets, individual models of population status, and ensemble methods to combine those estimates. We then describe how we evaluated the ability of the various models to estimate population status.

## Datasets
-->

We first developed and tested ensemble methods on a fully factorial simulated dataset [@rosenberg2014] with known status. These published simulated data included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of biomass depletion at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four exploitation settings: (1) a constant exploitation rate, (2) a exploitation rate coupled with biomass to mimic an open-access single-species fishery, (3) a scenario where exploitation rate increases continuously, and (4) a 'roller-coaster' scenario where the exploitation rate increases and then decreases. Process noise (recruitment variability; i.e. unexplained variability in population dynamics) was introduced to the models at two magnitudes in log space, $N(0, 0.2^2)$ and $N(0, 0.6^2)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with $N(0, 0.2^2)$ measurement error around log(catch) and one scenario without measurement error.

<!--
TODO: Jamie: add table instead / in addition to this paragraph?
-->

@rosenberg2014 ran ten iterations for each combination of factors adding stochastic draws of recruitment and catch-recording variability each time to generate a total of 5760 stocks. The simulation models were built in the package `FLR` [@kell2007] for the statistical software \textsf{R} [@r2015]. The code to generate the simulations is available at <https://github.com/iagomosqueira/stockStatusFAO>.

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database [@ricard2012] --- a compilation of stock-assessment output from hundreds of exploited marine populations around the world. Our analysis of the stock-assessment database was based on version XX downloaded on XX (TODO). After removing stocks for which the models described below did not converge (TODO: add number), this database included `r ram_stocks_n` stocks.
<!-- TODO Dan: Do we address the implications of only running our method on stocks that converged later on? i.e. is there something about those omitted stocks that would make them less suited for this type of analysis?-->

@@

\noindent 
where $L$ is a generalized loss function, $f$ is some generalized regression function, $\hat{b}_\mathrm{CMSY}$ is the mean or slope prediction from the CMSY model etc.\ (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the spectral density of the scaled catch series at frequencies of 0.20 and 0.05, $\hat{b}$ are the mean or slope values the model is fitted to, and $\hat{\theta}$ is the set of estimated parameters for the superensemble regression function. The $\argmin_\theta$ refers to finding the set of $\theta$ that minimize the loss function. For all superensemble models of mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero --- we fit the models in log space and exponentiated the predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit superensemble models on the natural untransformed scale.
<!--add theta-hat description, b hat, make sure these are right, maybe simplify, describe arg min-->
